---
title: "Project"
author: "Ritwik Bhriguvanshi"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}

library(MASS)
library(ggplot2)
library(grid)
library(gridExtra)
library(GGally)
library(leaps)
library(glmnet)
library(dplyr)

## Load the Dataset

FL = readRDS("FL_data.rds")
FL_data = readRDS("FL_data.rds")
colnames(FL_data)
head(FL_data)

str(FL_data)

## HAC is basically a Binary Variable 
## 1 if Hospital Acquired Condition has occured, 0 if not 

FL_data$HACs <- as.numeric(FL_data$HACs)
FL_data$HACs <- ifelse(FL_data$HACs == 1,0,1)
head(FL_data$HACs)
sum(FL_data$HACs)   ## [1] 72968

## No missing values
sum(is.na(FL_data$HACs))

FL_data$HACs <- as.factor(FL_data$HACs)

```

```{r Initial EDA}

corrmap <- FL[,c('age','nchronic','ndx','npr','totchg','los','count_dxnopoa_ns','visit','ahour','totalhospital_noofbed','HACs')]
corrmap$HACs <- as.numeric(as.character(corrmap$HACs))
heatmap(cor(corrmap), Rowv = FALSE, density.info = "none")

```

```{r pressure, echo=TRUE}

## Changing HAC to "Yes" or "No"
FL_data$HACs <- ifelse(FL_data$HACs == 0,"No", "Yes")

## Initial EDA -- Boxplot Analysis

par(mfrow=c(1,2))
boxplot (age ~ HACs,data = FL_data, ylab = "age", col = "blue")
boxplot (nchronic ~ HACs,data = FL_data, ylab = "nchronic", col = "yellow")
boxplot (ndx ~ HACs,data = FL_data, ylab = "ndx", col = "gray")
boxplot (los ~ HACs,data = FL_data, xlab = "los", col = "orange",horizontal = T)
boxplot (visit ~ HACs,data = FL_data, xlab = "visit", col = "lightcoral", horizontal = T)
boxplot (ahour ~ HACs,data = FL_data, ylab = "hour", col = "darkolivegreen1")

```

```{r Numerical Summary}

## Count of HAC on Heart Failure Patients
HF <- FL_data %>%
filter(., drg %in% c(291:293))%>%
group_by(., HACs) %>%
summarize(., 'Heart Failure Patients' = n())
print(HF)

## Count of AMI on Heart Attack Patients
AMI <- FL_data %>%
filter(., drg %in% c(280:282))%>%
group_by(., HACs) %>%
summarize(., 'Heart Attack Patients' = n())
print(AMI)

## Count of Pneumonia on Heart Attack Patients
PN <- FL_data %>%
filter(., drg %in% c(193:195))%>%
group_by(., HACs) %>%
summarize(., 'Pneumonia Patients' = n())
print(PN)

## Summary of all the Variables

FL2 <- FL_data
FL2$teachstatus <- as.numeric((FL2$teachstatus))
FL2$tcontrol <- as.numeric(as.character(FL2$tcontrol))
FL2$rural  <- as.numeric(as.character(FL2$rural))
FL2$pay1 <- as.numeric(as.character(FL2$pay1))
FL2$condition <- as.numeric((FL2$condition))
FL2$drg <- NULL #drg and condition are just the same thing
FL2$atype <- as.numeric(as.character(FL2$atype))
FL2$aweekend <- as.numeric(as.character(FL2$aweekend))
FL2$dqtr <- as.numeric(as.character(FL2$dqtr))
FL2$dshospid <- as.numeric(as.character(FL2$dshospid))
FL2$female <- as.numeric(as.character(FL2$female))
FL2$medincstq <- as.numeric(as.character(FL2$medincstq))
FL2$race <- as.numeric(as.character(FL2$race))
FL2$year <- as.numeric(as.character(FL2$year))
FL2$tran_out <- as.numeric(as.character(FL2$tran_out))
FL2$zipinc_qrtl <- as.numeric(as.character(FL2$zipinc_qrtl))
FL2$cm_anemdef <- as.numeric(as.character(FL2$cm_anemdef))
FL2$cm_chf <- as.numeric(as.character(FL2$cm_chf))
FL2$cm_chrnlung <- as.numeric(as.character(FL2$cm_chrnlung))
FL2$cm_coag <- as.numeric(as.character(FL2$cm_coag))
FL2$cm_depress <- as.numeric(as.character(FL2$cm_depress))
FL2$cm_dm <- as.numeric(as.character(FL2$cm_dm))
FL2$cm_dmcx <- as.numeric(as.character(FL2$cm_dmcx))
FL2$cm_htn_c <- as.numeric(as.character(FL2$cm_htn_c))
FL2$cm_hypothy <- as.numeric(as.character(FL2$cm_hypothy))
FL2$cm_lytes <- as.numeric(as.character(FL2$cm_lytes))
FL2$cm_neuro <- as.numeric(as.character(FL2$cm_neuro))
FL2$cm_obese <- as.numeric(as.character(FL2$cm_obese))
FL2$cm_perivasc <- as.numeric(as.character(FL2$cm_perivasc))
FL2$cm_renlfail <- as.numeric(as.character(FL2$cm_renlfail))
FL2$read <- as.numeric((FL2$read))

cbind(FL2 %>% group_by(HACs)%>% 
summarise_all(funs(mean(., na.rm = TRUE))), count = c(table(FL2$HACs)))

```

Count for Heart Failure Patients (With HAC = 1): 33909
Count for Heart Attack Patients (With HAC = 1): 11733	
Count for Heart Pneumonia Patients (With HAC = 1): 27326	

Other interesting observations:
HAC 1 is attributed to people with higher age, about 75 years of age.
HAC 1 cases found more in chronic diseases.
Higher number of ICD-9 discharge and procedures attributing to HAC 1
LOS is significantly higher for HAC = 1 vs HAC = 0; almost twice.
HAC 1 cases are attributed to small size of hospitals with less number of beds.
As mentioned earlier, Comorbidity is prevelant with HAC1 cases. 

Other numerical variables such as hour and visit is the same in either case.

```{r Contigency Tables}

table(FL_data$cm_anemdef, FL_data$HACs)
table(FL_data$cm_chf, FL_data$HACs)
table(FL_data$cm_chrnlung, FL_data$HACs)
table(FL_data$cm_coag, FL_data$HACs)
table(FL_data$cm_depress, FL_data$HACs)
table(FL_data$cm_dm, FL_data$HACs)
table(FL_data$cm_dmcx, FL_data$HACs)
table(FL_data$cm_htn_c , FL_data$HACs)
table(FL_data$cm_hypothy, FL_data$HACs)
table(FL_data$cm_lytes, FL_data$HACs)
table(FL_data$cm_neuro , FL_data$HACs)
table(FL_data$cm_obese, FL_data$HACs)
table(FL_data$cm_perivasc , FL_data$HACs)
table(FL_data$cm_renlfail, FL_data$HACs)

```

Each of these are some type of different disorders; coded as a binary variable. Interesting observation here is that - the number of HAC = 1 prevelence is less when only one of these disorders is taken into consideration, but logically not in many cases patients will have combination of these disorders. Taking that into consideration, a handful of HAC = 1 cases exists in each of these different disorders.

```{r EDA-LOS}

## Plotting LOS and log LOS against HACs
p1 = ggplot(FL_data, aes(x=los, color = factor(HACs))) + geom_histogram(bins = 15) + theme_bw()
p2 = ggplot(FL_data, aes(x=log(los), color = factor(HACs))) + geom_histogram(bins = 15) + theme_bw()

gridExtra::grid.arrange(p1,p2,ncol=2)

## LOS By Year
boxplot(los ~ HACs*year,data = FL_data, ylab = "year", col = "lightcoral", las = 2, horizontal = T, main = "LOS vs Year, differentiated by HACs")

## LOS by Three different conditions overlayed by HAC
ggplot(data=FL_data, aes(x=los,y=condition, colour=HACs)) + geom_point()

```


```{r EDA-TotChg}

## Plotting TotChg and log TotChg against HACs
p3 = ggplot(FL_data, aes(x=totchg, color = factor(HACs))) + geom_histogram(bins = 15) + theme_bw()
p4 = ggplot(FL_data, aes(x=log(totchg), color = factor(HACs))) + geom_histogram(bins = 15) + theme_bw() ## Log Transformation makes it normal

gridExtra::grid.arrange(p3,p4,ncol=2)

## totchg By Year
boxplot(totchg ~ HACs*year,data = FL_data, ylab = "year", col = "lightcoral", las = 2, horizontal = T, main = "totchg vs Year, differentiated by HACs")

## totchg by Three different conditions overlayed by HAC
ggplot(data=FL_data, aes(x=totchg,y=condition, colour=HACs)) + geom_point()
## Blue Points which have HAC == 1 are towards the higher end (right corner)

```


