---
title: "Homework 1"
Author: Ritwik Bhriguvanshi
Professor: Dr. Hilafu
Class: BZAN 540
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

FL_2014_CORE <- readRDS("FL_2014_CORE.rds")
library(data.table)
library(dplyr)
library(ggplot2)

names(FL_2014_CORE) <- tolower(names(FL_2014_CORE))

FL_2014 <- select(FL_2014_CORE,
                  age,atype,aweekend,died,drg,female,dqtr,dshospid,
                  los,nchronic,ndx,npr,pl_cbsa,race,visitlink,mdc,
                  zip,totchg,orproc,pay1)

FL_2014 <- FL_2014 %>% 
  mutate(los = as.integer(los), totchg = as.integer(totchg), atype = as.factor(atype),
         dqtr = as.factor(dqtr), aweekend = as.factor(aweekend), pay1 = as.factor(pay1))
```


```{question 1a}
q1 <- FL_2014 %>%
select(died, drg, dshospid)%>%
filter(., drg %in% 291:293) %>%     #Filter for heart failure patients
group_by(., dshospid) %>%
summarize(., Count = n(),TotalDeath = sum(died), MortalityRate = mean(died))%>%
arrange(desc(MortalityRate))

# Printing top 10 Hospital IDs in terms of Highest Mortality Rate
q1 %>%
top_n(10,MortalityRate)

```
The histogram is heavily skewed towards the right because with the increase in mortality rate,
the number of hospitals are decreasing. There are a large number of hospitals who have mortality rate between 0.0 and 0.1 for heart failure hospitalizations.

```{r question 1a plot}
Graph0 <- ggplot(q1,aes(x=MortalityRate)) + 
  geom_histogram(binwidth = 0.01,col="black", 
                 fill="red", 
                 alpha = .2) 

Graph0 + theme_bw()+
  labs(title="Mortality Rate for Heart Failure Hospitalizations") +
  labs(x="Mortality Rate", y="Number of Hospitals",
       caption = " Plot for question 1a: made by RB") 
```
The histogram is heavily skewed towards the right because with the increase in mortality rate,
the number of hospitals are decreasing. There are a large number of hospitals who have mortality rate between 0.0 and 0.1 for heart failure hospitalizations.

```{r question 1b plot1}
mygraph1 <- ggplot(q1, aes(x = q1$Count, y = q1$MortalityRate)) +
  geom_point()    

mygraph1 +
  theme_light() +
  labs(
    x = "Total Number of Hospitalization",
    y = "Mortality Rate",
    color = "Green",
    title = "Association between Hospital Mortality Rate and Total Number Hospitalizations",
    subtitle = "For Heart Failure Patients grouped by Hospital ID",
    caption = "Plot 1 for question 1b: made by RB")
```
The mortality rate is high initially is it depends on the number of admissions. For 
example, if we have 2 admits in a hospital and one of them dies, the mortality rate will be
considerably high. This is primary the reason why we see high mortality rates with low 
number of hospital admits. As we move forward in the number of hospitalization, the mortality
rate considerably drops down below 0.1

```{r question 1b code}
q2 <- FL_2014 %>%
  select(died, drg, dshospid)%>%
  group_by(., dshospid) %>%
  summarize(., Count2 = n(),TotalDeath2 = sum(died), MortalityRate2 = mean(died))%>%
  arrange(desc(MortalityRate2))

Merged <- merge(q1,q2)

```

```{r question 1b plot2}
mygraph2 <- ggplot(Merged, aes(x = Merged$Count2, y = Merged$MortalityRate2)) +
  geom_point()    

mygraph2 +
  theme_light() +
  labs(
    x = "Total number of Hospitalizations",
    y = "Mortality Rate for all patients",
    color = "Green",
    title = "Association on Mortality Rate",
    subtitle = "For Heart Failure Patients and all the Patients",
    caption = "Plot 2 for question 1b: made by RB")
```

The plot doesn't provide sufficient evidence for association between mortality rate 
for all the patients and 'only' heart failure patients. Majority of the data points are condensed near the origin depitcing in addition to the correlation being 0.44 which is a weak correlation.

```{r question 1c}
q3 <- FL_2014 %>%
  select(died, drg, dshospid,totchg,los)%>%
  filter(., drg %in% 291:293) %>%     #Filter for heart failure patients
  group_by(., dshospid) %>%
  summarize(., Count = n(),TotalDeath = sum(died), MortalityRate = mean(died),
            LOS = sum(los), TotalCharge = sum(totchg) , AverageDailyCharge = sum(totchg)/sum(los))%>%
  arrange(desc(MortalityRate))

```

```{r question 1c plot}
mygraph3 <- ggplot(q3, aes(y = q3$TotalCharge, x = q3$LOS)) +
  geom_point()    

mygraph3 +
  theme_light() +
  labs(
    x = "Length of Stay in Hospital",
    y = "Total Charge",
    color = "Green",
    title = "Association between Total Charge and Length of Stay",
    subtitle = "For Only Heart Failure Patients",
    caption = "Plot 1 for question 1c: made by RB") +
  theme(plot.title = element_text(size = 8, face ='bold'))
```

There is a strong association between length of stay in hospital and total charge a
associated with it. As the duration of stay increases, the total cost of treatment increases.
These two variables have a positive relationship. Additionally, total charge is dependent 
on the length of stay.


```{r question 1c plot2}
mygraph4 <- ggplot(q3, aes(y = q3$AverageDailyCharge, x = q3$Count)) +
  geom_point()   

mygraph4 +
  theme_light() +
  labs(
    x = "Total Number of Admits",
    y = "Average Daily Charge",
    color = "Green",
    title = "Association between Average Daily Charge and Number of Hospitalizations",
    subtitle = "For Only Heart Failure Patients",
    caption = "Plot 2 for question 1c: made by RB") +
    theme(plot.title = element_text(size = 8, face ='bold'))

```

Weak association between Average Daily Charge and Number of Total Admits in the 
hospital. There is not definite pattern (increase or decrease) with the increase in the 
number of admits.

```{r question 1d}
q4 <- FL_2014 %>%
  select(drg, dshospid, dqtr, totchg,)%>%
  filter(drg %in% 291:293)%>%            # Filter for heart failure patients
  group_by(dqtr)%>%
  summarise(N = n(), TotalCharges = mean(totchg))

```

```{r question 1d plot}
mygraph5 <- ggplot(q4, aes(y= N, x= q4$dqtr))+
  geom_bar(stat="identity", fill="green")

mygraph5 + ggtitle("Total Number of Hospitalization by Quarter") +
  xlab("Quarter of the year") + ylab("Number of Hospitalizations")+
        labs(caption = "Plot 1 for question 1d: made by RB")+
    theme_light()
```

The number of hospital admits stays in the same range for all the quarters. 
It marginally varies between the four quarters. It is highest in the first quarter and lowest
in the third quarter.

```{r question 1e}
q5 <- FL_2014 %>%
    select(drg, dshospid, dqtr, totchg, pay1, los)%>%
    filter(drg %in% 291:293)%>%            # Filter for heart failure patients
    group_by(pay1)%>%
    mutate(DailyCharges = as.numeric(totchg) / as.numeric(los))

# 1	Medicare
# 2 Medicaid
# 3	Private insurance
# 4	Self-pay
# 5	No charge
# 6	Other

```

```{r question 1e plot1}
mygraph6a <- ggplot(q5, aes(x= pay1,y= los))+
  geom_boxplot(aes(group=pay1),outlier.shape = NA) + scale_y_continuous(limit = c(0,15))

mygraph6a + ggtitle("Length of Stay and Type of Pay") + 
  xlab("Type of Pay") + ylab("Length of Stay") + 
  labs(caption = "Plot 2 for question 1e: made by RB") +
  theme_light()

```

The length of stay has a median of less than five regardless of its pay type.
The boxplots display a similar range and does not vary much amonth the different pay types.
1559 observations were ommited after setting the higher sclae of 15 on the Y-axis.



```{r question 1e plot2}
mygraph6 <- ggplot(q5, aes(x= pay1,y= DailyCharges))+
  geom_boxplot(aes(group=pay1),outlier.shape = NA) + scale_y_continuous(limit = c(0,25000))

mygraph6 + ggtitle("Daily Charges and Type of Pay") + 
xlab("Type of Pay") + ylab("Daily Charges") + 
  labs(caption = "Plot 1 for question 1e: made by RB") +
theme_light()

```

The median for daily charges is around $7500 regardless of its pay type.
There is not much variation among the different boxplots. 
2417 rows were ommitted after setting an upper limit of 25000 on the Y-axis.

```{r question 1f}

q6 <- FL_2014 %>%
  select(died, drg, dshospid, aweekend)%>%
  filter(., drg %in% 291:293) %>%     #Filter for heart failure patients
  group_by(., dshospid, aweekend) %>%
  summarize(., Count = n(),TotalDeath = sum(died), MortalityRate = mean(died))%>%
  arrange(desc(MortalityRate))

# '0'	Admitted Monday-Friday
# '1'	Admitted Saturday-Sunday

```

```{r question 1f plot}
mygraph7 <- ggplot(q6, aes(x = aweekend, y = MortalityRate, group = aweekend)) +
  geom_boxplot() + scale_y_continuous(limit = c(0,.1)) +
  theme_light()
  
  mygraph7 + ggtitle("Mortality Rate and Weekend Analysis") + 
  xlab("Day of the week (0 for weekday, 1 for weekend)") + ylab("Mortality Rate") +
    labs(caption = "Plot 1 for question 1f: made by RB")
  
```

There is no significant different on mortality rate based on the day of the week.
The median for both the cases is above 1.25% but less than 2.5%.
```{r question 1g }
q7 <- FL_2014 %>%
  select(died, drg, dshospid, atype, los)%>%
  filter(., drg %in% 291:293) %>%     #Filter for heart failure patients
  group_by(., dshospid)
```

```{r question 1g plot}
mygraph8 <- ggplot(q7, aes(x = atype, y = los, group = atype)) +
  geom_boxplot() + scale_y_continuous(limit = c(0,10)) +
  theme_light()

mygraph8 + ggtitle("Mortality Rate and Weekend Analysis") + 
  xlab("Type of Treatment") + ylab("Length of Stay") +
  labs(caption = "Plot 1 for question 1g: made by RB")

```

The median number for length of stay is higher for type 3 (elective) and type 5(Delivery).
There is no length of stay of newborn (type 4). The hiher limit for length of stay on the 
Y-axis was set at 10, which ommitted 4584 rows.

```{r question 1g Linear Model}

linearMod <- lm(los ~ atype, data=q7)
linearMod
summary(linearMod)

```

Only 0.2% of the variation is explained by the variable 'atype'. Hence, the model does not 
provide a good estimate for the length of stay. 

```{r question 1h }
q8 <- FL_2014 %>%
  select(died, drg, dshospid, totchg, los)%>%
  filter(., drg %in% 291:293) %>%     #Filter for heart failure patients
  group_by(., dshospid)

```

```{r question 1h plot}
mygraph9 <- ggplot(q8, aes(x = los, y = totchg, group = los)) +
  geom_point() +
  theme_light() 

mygraph9 + ggtitle("Total Charge vs Length of Stay") + 
  xlab("Length of Stay") + ylab("Total Charges")+
  labs(caption = "Plot 1 for question 1h: made by RB")
```


```{r question 1h Linear Model}
linearMod <- lm(totchg ~ los, data=q8)
linearMod
summary(linearMod)

```

Slope is 6769.71 - An increase in stay at the hospital by each additional day 
will increase the cost by $6770. 
R-sqaure is approximately 60% which is extremely good, as one variable out of the 302 
variables in the data explain about 60% variation in the model.
RMSE is 23660, which portrays the typical error when we predict the total charge on
the basis of length of stay.

```{r question 2}
data <- read.csv("HW1PROB8.csv")
data <- data[-c(1:21,82:102),]

```

```{r question 2a}
lm1 <- lm(total_quantity ~ avg_price, data = data)
lm1
summary(lm1)
plot(lm1)

```

```{r question 2b}
(leverage1 <- hatvalues(lm1))
(StanRes1 <- rstandard(lm1))
(residual1 <- lm1$residuals)

```

```{r question 2b plot}
(plot <- plot(data$week_no, StanRes1))
```

```{r question 2c}
predict(lm1, newdata = (avg_price= 1.59),interval="confidence",level=0.95)
```

```{r question 2c plot}
(plot <- plot(data$avg_price, StanRes1))

```

```{r question 2d}
lm2 <- lm(log(total_quantity) ~ log(avg_price), data = data)
lm2
summary(lm2)

predict(lm2, newdata = (avg_price= 1.59),interval="confidence",level=0.95)

```
