---
title: "Infographics"
author: "Ritwik Bhriguvanshi"
date: "11/9/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
data <- read.csv("adult.csv", header = T, na.strings = c("", "NA", "?", "-"))
dim(data)
head(data)
```

Step1: Explore the data, change the type, analyze tables

```{r basicexploration}
## Perform Exploratory data Analysis
str(data)

## Drop Two Columns "fnlwgt" and "educational.num
drop <- c("fnlwgt","educational.num")
data = data[,!(names(data) %in% drop)]

## Missing Values for Each Column
colSums(is.na(data))

## Change Type of Variables
data$workclass <- as.character(as.factor(data$workclass))
data$education <- as.character(as.factor(data$education))
data$marital.status <- as.character(as.factor(data$marital.status))
data$occupation <- as.character(as.factor(data$occupation))
data$relationship <- as.character(as.factor(data$relationship))
data$race <- as.character(as.factor(data$race))
data$gender <- as.character(as.factor(data$gender))
data$native.country <- as.character(as.factor(data$native.country))
data$income  <- as.character(as.factor(data$income))

## Change NA values to Not Known
data$workclass[is.na(data$workclass)] <- "Not Known"
data$occupation[is.na(data$occupation)] <- "Not Known"
data$native.country[is.na(data$native.country)] <- "Not Known"

## Make Table for Categorical Data
## Two of these tables have been used in the Poster for Analysis

table(data$workclass) # Missing values exist
table(data$education)
table(data$marital.status)
table(data$occupation) # Missing values exist
table(data$relationship)
table(data$race)
table(data$gender)     
table(data$native.country) # Missing values exist
table(data$income)

```

Step2: Boxplot Analysis for relevant numerical variables

```{r boxplot}
## Visualize Income with Categorical Data
par(mfrow=c(1,2))
boxplot(age ~ income,data = data, ylab = "age", col = "cyan3")
boxplot(hours.per.week ~ income,data = data, ylab = "Working Hours per Week",xlab = "Income Level", col = "darkkhaki")
boxplot(capital.gain ~ income,data = data, ylab = "Capital Gain", col = "lightcoral")
boxplot(capital.loss ~ income,data = data, ylab = "Capital Loss", col = "lightgreen")
```

Step3: In-Depth Analysis for "Age" variable

```{r ageanalysis}
library(ggplot2)
# See Distribution of Age
p <- ggplot(data, aes(x=age)) + geom_histogram(binwidth=1,color="black", fill="lightblue")
## Add mean Line for Age
p + geom_vline(aes(xintercept=mean(age)),
            color="black", linetype="dashed", size=1)

# grouped by number of gears (indicated by color)
qplot(age, data=data, geom="density", fill=income, alpha=I(.65),
   main="Distribution of Age", xlab="Age",
   ylab="Density")

## Make 5 levels from the Age group to make a new column
data$agegroups <- cut(data$age,b=5)

## Distribution of Age Group in the five cuts made
table(data$agegroups)
data$agegroups <- as.numeric(as.factor(data$agegroups))

# boxplot(agegroups ~ income,data = data, ylab = "age", col = "yellow")

data$agegroups[data$agegroups == 1] <- "16.9,31.6"
data$agegroups[data$agegroups == 2] <- "31.6,46.2"
data$agegroups[data$agegroups == 3] <- "46.2,60.8"
data$agegroups[data$agegroups == 4] <- "60.8,75.4"
data$agegroups[data$agegroups == 5] <- "75.4,90.1"

## Distribution of Age Group in the five cuts made
table(data$agegroups)

# grouped by number of gears (indicated by color)
qplot(agegroups, data=data, geom="density", fill=income, alpha=I(.5),
   main="Distribution of Age", xlab="Age",
   ylab="Density")

```

Step4: Analyze each dimension in the dataset by group it by a variable & income

```{r dplyr}
colnames(data)
library(dplyr)
## Workclass
workclass <- data %>%
group_by(workclass, income) %>%
summarize("Count" = n(), "Mean of Age" = round(mean(age),2), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(workclass)

# Education
education <- data %>%
group_by(education, income) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(education)

# Martial Status
martial <- data %>%
group_by(marital.status, income) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(martial)

## Occupation
occupation <- data %>%
group_by(occupation, income) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(occupation)

## Relationship
relationship <- data %>%
group_by(relationship, income) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(relationship)

## Race
race <- data %>%
group_by(race, income) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(race)

## Gender
gender <- data %>%
group_by(gender, income) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(gender)

## Native Country
country <- data %>%
group_by(native.country) %>%
summarize("Count" = n(), "Percentage" = n()/nrow(data)*100) %>%
arrange(desc(Count))
print(country)

```

Step5: Analyze each variable when Income is greater than 50K

```{r visualmorethan50}
above50k <- data %>%
  filter(income == ">50K") 

## Age Graph
age_above50k <- ggplot(above50k , aes(x=age)) + geom_histogram(binwidth=1,color="black", fill="lightblue")                      + labs(title="Distribution of Age",x="Age", y = "Frequency")
print(age_above50k)

## Age Stat
age_above_50k_stat <- above50k %>%
                      summarise("Average Age" = mean(age), "Median Age" = median(age))
print(age_above_50k_stat)

## Workclass Table
workclass_above50k <- above50k %>%
                group_by(workclass) %>%
                summarise("Count" = n(), Percentage = n()/nrow(above50k)*100) %>%
                arrange(desc(Count)) %>%
                top_n(6)
print(workclass_above50k)

## Education Distribution
edu_above50k <- above50k %>%
              group_by(education) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Count)) %>%
              top_n(5)
print(edu_above50k)

## Marital Status
marital_above50k <- above50k %>%
              group_by(marital.status) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Percentage))
print(marital_above50k)

## Occupation
occupation_above50k <- above50k %>%
              group_by(occupation) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Percentage)) %>%
              top_n(6)
print(occupation_above50k)

## Relationship
relationship_above50k <- above50k %>%
              group_by(relationship) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Percentage))
print(relationship_above50k)

## Race
race_above50k <- above50k %>%
              group_by(race) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Percentage))
print(race_above50k)

## Gender
gender_above50k <- above50k %>%
              group_by(gender) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Percentage))
print(gender_above50k)

## Capital Gain
capitalgain_above50k <- above50k %>%
              summarise("Mean Capital Gain" = mean(capital.gain), "Minimum Capital Gain" = 
                           min(capital.gain), "Maximum Capital Gain" = max(capital.gain))
print(capitalgain_above50k)

## Capital Loss
capitalloss_above50k <- above50k %>%
              summarise("Mean Capital Loss" = mean(capital.loss), "Minimum Capital Loss" = 
                           min(capital.loss), "Maximum Capital Loss" = max(capital.loss))
print(capitalloss_above50k)

## Hours per week
hours_above50k <- above50k %>%
              summarise("Mean Hours per week"= mean(hours.per.week),"Median Hours per week"= median(hours.per.week)) %>%
              arrange(desc("Mean Hours per week"))
print(hours_above50k)

## Native Country
country_above50k <- above50k %>%
              group_by(native.country) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(above50k)*100) %>%
              arrange(desc(Percentage)) %>%
              top_n(5)
print(country_above50k)
```

Step6: Analyze each variable when income is less than 50K

```{r visuallessthan50}
below50k <- data %>%
  filter(income == "<=50K")  

## Graph
age_below_50k <- ggplot(below50k , aes(x=age)) + geom_histogram(binwidth=1,color="black", fill="lightblue") + labs(title="Distribution of Age",x="Age", y = "Frequency")
print(age_below_50k)

## Age Stat
age_below_50k_stat <- data %>%
                       summarise("Average Age" = mean(age), "Median Age" = median(age))
print(age_below_50k_stat)

## Workclass Table
workclass_below50k <- below50k %>%
                group_by(workclass) %>%
                summarise("Count" = n(), Percentage = n()/nrow(below50k)*100) %>%
                arrange(desc(Count)) %>%
                top_n(6)
print(workclass_below50k)

## Education Distribution
edu_below50k <- below50k %>%
              group_by(education) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Count)) %>%
              top_n(6)
print(edu_below50k)

## Marital Status
marital_below50k <- below50k %>%
              group_by(marital.status) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Percentage))
print(marital_below50k)

## Occupation
occupation_below50k <- below50k %>%
              group_by(occupation) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Percentage)) %>%
              top_n(6)
print(occupation_below50k)

## Relationship
relationship_below50k <- below50k %>%
              group_by(relationship) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Percentage))
print(relationship_below50k)

## Race
race_below50k <- below50k %>%
              group_by(race) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Percentage))
print(race_below50k)

## Gender
gender_below50k <- below50k %>%
              group_by(gender) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Percentage))
print(gender_below50k)

## Capital Gain
capitalgain_below50k <- below50k %>%
              summarise("Mean Capital Gain" = mean(capital.gain), "Minimum Capital Gain" = 
                           min(capital.gain), "Maximum Capital Gain" = max(capital.gain))
print(capitalgain_below50k)

## Capital Loss
capitalloss_below50k <- below50k %>%
              summarise("Mean Capital Loss" = mean(capital.loss), "Minimum Capital Loss" = 
                           min(capital.loss), "Maximum Capital Loss" = max(capital.loss))
print(capitalloss_below50k)

## Hours per week
hours_below50k <- below50k %>%
              summarise("Mean Hours per week"= mean(hours.per.week),"Median Hours per week"= median(hours.per.week)) %>%
              arrange(desc("Mean Hours per week"))
print(hours_below50k)

## Native Country
country_below50k <- below50k %>%
              group_by(native.country) %>%
              summarise("Count"= n(), "Percentage" = n()/nrow(below50k)*100) %>%
              arrange(desc(Percentage)) %>%
              top_n(6)
print(country_below50k)
```

Step7: Some other variable Interactions and Visualizations

```{r plot}
## Some Interesting Interactions between variables 
par(mfrow=c(1,2))
boxplot (age ~ gender*income,data = data, ylab = "age", col = "darkkhaki", cex.axis=.55)
boxplot (age ~ race*income,data = data, ylab = "age", col = "green", cex.axis = .4)

##################################   Pie- Chart   ##########################################
total_gender <- table(data$gender)
pie(total_gender, labels=as.character(rownames(total_gender)), col=c("lightblue","lightgreen", border="black", clockwise=TRUE))
print(total_gender)

total_race <- table(data$race)
pie(total_race, labels=as.character(rownames(total_race)), col=c("lightblue","lightgreen", "pink","orange","grey", border="black", clockwise=TRUE))
print(total_race)

##################################   Plotly    ##########################################
library(plotly)
## Occupation for Below 50K
fig <- plot_ly(
  type = "funnelarea",
  text = occupation_below50k$occupation,
  values = occupation_below50k$Percentage)
fig

## Occupation for Above 50K
fig1 <- plot_ly(
  type = "funnelarea",
  text = occupation_above50k$occupation,
  values = occupation_above50k$Percentage)
fig1

```





