---
title: "HW4"
author: "Ritwik Bhriguvanshi"
date: "10/14/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r 4.12}
# Q 4.12

# a. Based on the parameter estimates, which group is most likely to have the “yes” response? Estimate the probability in that case.

intercept = -3.5961
coeff_def = -0.8678
coeff_vic = 2.4044


(exp(coeff_def))
(exp(coeff_vic))

#based on the parameters estimates, for a given defendant's race  the group outcome likely to to have a death penalty is when the victim is white with estimated odds of exp(2.4044)= 11 times the estimated odds when the victim is black. 

#based on the parameters estimates, for a given victim's race the group outcome likely to to have a death penalty is when the defendant is white with estimated odds of exp(-0.8678)= 0.41 times less the estimated odds when the defendant is white. we can conclude that black defendants would receive a yes respose or death penalty 

# estimated the probability = odds/(1+odds) when the victim is white

a <- -3.5961+2.4044
(exp(a)/(1+exp(a)))
# 0.232955

b <- 2.4044 + 0.6006*11
(exp(b)/(1+exp(b)))
# 0.999878


# b. Interpret the parameter estimate for victim’s race.

#The estimated odds to receive a death penalty given that its a white victim is exp(2.4044) = 11 times the estimated odds when the victim is a black victim

#The estimated odds to receive a death penalty given that the defendant is white defendant is exp(−0.8678)=0.41 times less the estimated odds when the defendant is black


# c. Using information shown, construct and interpret a 95% likelihood-ratio confidence interval for the conditional odds ratio between the death penalty verdict and victim’s race.

# 95% likelihood confidence interval for conditional odds is (3.694333 41.161362)
c(exp(1.3068),exp(3.7175))
#the true estimated odd of a dealth penalty when the vicitim is white lie within the range of (3.694333 41.161362) times compared to when the victim is Black 


# d. Test the effect of victim’s race, controlling for defendant’s race, using a Wald test or likelihood-ratio test. Interpret.

#The effect of Victim's race by controlling the defendant race using a wald's test
#H0: B =0
#H1: B /=0

#where B is the coeff of victim's race

(Wald_statistics = (2.4044/0.6006))
#[1] 4.00333 > 1.95

# Reject Null, the effect of victim's race controlling for defendant's race is not independent. SOme effect of victim's race plays a role in the penalty outcome.

```


```{r q2}
# Q.4.16

Ex4.16 <- read.csv ("Ex4.16.csv", sep = ",", header = T)
str(Ex4.16)

### Change column "tf" into Categorical instead of Boolean 
Ex4.16$tf <- as.character(Ex4.16$tf)
### Use ifelse statement to Change into 'T' or 'F'
Ex4.16$tf <- ifelse(Ex4.16$tf == "TRUE", "T", "F")
## Check
head(Ex4.16)

## Build the Model
model <- glm(cbind(Yes, No)~factor(tf) + factor(jp) + factor(EI) + factor(SN), data=Ex4.16, family = "binomial")
summary(model)

## For finding pie for ESTJ, this is the first row in the list, we will use this to make prediction
predict(model, type="response")[1]

## For finding pie for ENTJ, this is the first row in the list, we will use this to make prediction
predict(model, type="response")[6]
```

Response:
(a) The predictor variables were coded as factors.
(b) ESTJ: 0.135186 
(c) Highest possible set of coefficients, which makes personality type of ENTP as the highest. The estimated probability will be around 0.2271486 

```{r q3}
## Part(a)
## Value of Pie (Use the intercept from the given Model Output)
(pie <- exp(−2.8291)/(1+exp(−2.8291)))

## Part(b)
## Conditional odds Ratio
(odds = exp(0.5805))

## Part (C)
## 95% likelihood-ratio confidence interval for the conditional odds ratio
cbind(exp(0.1589), exp(1.008))

## Part (d)
## For E/I scale, the estimated conditional odds ratio b/w E/I and Response variable is:
(conditional_odds = (1/(exp(0.5805))))
(CI <- cbind((1/exp(1.008)), (1/exp(0.1589))))

```

Response (a) : Value of Pie is 0.05577177

Response (b) : Given the T/F category, the estimated odds an extroverted person drinking frequently is about 1.786932 times the estimated odds an introverted person drinking frequently

Response (c) : Given the T/F category, there is a 95% probability that the true odds ratio would likely be in the range [1.172221, 2.740115] for an extroverted person drinking frequently as compared to an introverted person drinking frequently.

Response (d) : There is a 95% condfidence that the conditional odds ratio between E/I and the response will exist within the range [0.3649481, 0.8530817], given that 'I' is used instead of E for the first category of E/I scale.

Response (e) : Let the parameter for E/I be denote by Beta. Let the null hypothesis be Beta = 0 and alternative hypothesis be Beta != 0. Looking at the Likelihood-ratio statistics value of 7.28 for E/I with 1 df and P-value of 0.007, we reject the null hypothesis since it is below the 5% threshold. Thus, we conclude that E/I has an effect on the response.

```{r q4}
#Q 4.30

data<-read.csv("NCAA Final.csv", header=TRUE,stringsAsFactors = TRUE)
#str(data)
model=glm(Count~Athlete + Graduate, family = poisson,data=data)
(summary(model))

(confint(model))
(exp(1.52940))
(exp(1.71675))
(exp(2.43042))

```

Our baseline/reference group for the model is Black females with no graduation
The deviance is 180 with 3 degrees of freedom. 
The chisquare value for the same is approximately close to zero, hence this model is not a good fit.

Conditional odd ratio for each athlete group is given below.
The estimated  odds of graduation given that the student is black male is exp(1.52904) = 4.62 times
The estimated  odds of graduation given that the student is white female is exp(1.71675) = 5.66 times
The estimated  odds of graduation given that the student is white male is exp(2.43042) = 11.36 times

The Confidence Interval has been presented in the code above.